---
title: "Motor Trend Data Analysis"
author: "Dinh Tuan Phan"
date: "5/3/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

## Synopsis

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

1. “Is an automatic or manual transmission better for MPG”
2. "Quantify the MPG difference between automatic and manual transmissions"  

## Install packages

```{r}
library(ggplot2)
```

## Load data

```{r}
data(mtcars)
head(mtcars)
```


Summary on data to understand the predictors and outcome

```{r}
summary(mtcars)
```


## First visualise the data

Run the pair plots to have a first glance on the correlation between input.

```{r}
pairs(mtcars, panel=panel.smooth, main="Relationship between predictors")

```


## Run statistical inference

We first run the correlation between the MPG and other predictors.

```{r}
cor(mtcars$mpg,mtcars[,-1])

```

We can see that the transmission type has a positive correlation (0.5998), so that answers the first question: “Is an automatic or manual transmission better for MPG”. The manual is better than automatic in term of mpg. We can check by plotting the am vs mpg.


## Plots

```{r}
ggplot(mtcars, aes(y=mpg, x=factor(am, labels = c("automatic", "manual")), fill=factor(am)))+
        geom_boxplot(colour="black", size=1)+
        xlab("Transmission Type") + ylab("Miles per Gallon (MPG)") + 
        scale_fill_discrete(name = "Transmission Type", labels = c("automatic", "manual"))
```


## Run regression model

We run linear model between all predictors vs outcome (mpg) and we examine parameters in the model.

```{r}
fullModel <-lm(mpg~.,mtcars)
summary(fullModel)
```

From the results, we can see that only wt is significant to the model because p-value < 0.1 (theoretically, p-values should be smaller than 0.05). We run the model again with only 1 predictor wt.


```{r}
lm1 <-lm(mpg~wt,mtcars)
summary(lm1)
```

This is a good model with very low p-value (0.001). It makes sense because the car efficiency (i.e., measured in mpg) heavily depending on its weight. We can further visualize by plotting.

```{r}
mtcars$tran=ifelse(mtcars$am == 0,"automatic","manual")
ggplot(mtcars, aes(x=wt, y=mpg, group=tran, color=tran, height=3, width=3)) + geom_point() +  
scale_colour_discrete(name = "Transmission Type",labels=c("Automatic", "Manual")) + 
xlab("Weight") + ggtitle("Weight vs MPG by Transmission Type")
```



Next we can try to run the model with one more predictor (i.e., the Transmission am) because predictor am has the lowest p-value among the remaining predictors.

```{r}
lm2 <-lm(mpg~wt+am,mtcars)
summary(lm2)
```

The predictor wt is significant (p-value less than 0.001) but am is not significant (p-value is to high 0.988) in the model.


Next, we might want to run a linear model with only predictor am to exam it independently.

```{r}
lm3 <-lm(mpg~am,mtcars)
summary(lm3)
```

This model seems to answer the question. The car efficiency (i.e., measured in mpg) depends on the transmission type.

The answer for the second question: "Quantify the MPG difference between automatic and manual transmissions" is: the cars with manual transmission is 7.245 mpg better than the automatic cars.


## Best model

```{r}
bestmodel = step(lm(data = mtcars, mpg ~ .), trace=0)
summary(bestmodel)
```
### Residual Analysis

We observe that the residuals seems randomly scattered around zero and that quantiles for the residuals fall somewhat close to the theoretical normal quantiles in the Q-Q plot.

```{r}
par(mfrow = c(2,2))
plot(bestmodel)
```



